---
title: "Variance of a Mixture of distributions"
author: "Marco Gramatica"
date: "1 August 2018"
header-includes:
   - \usepackage{hyperref}
output: pdf_document
---

# Formula

```{r, echo = F, message = F}
formula_var <- function(pop_d, mu_domains, sd_domains){
    pop_d <- pop_d/sum(pop_d)
  result <- pop_d%*%(sd_domains^2) + 
    (pop_d%*%(mu_domains^2) - (pop_d%*%mu_domains)^2)
  return(as.numeric(result))
}

# Domain order:
# a - b - c - ab - bc - ac - abc

frame_sd <- function(pop_d, mu_domains, sd_domains){
  pos_select <- list(c(1, 4, 6:7), # Frame A
                     c(2, 4:5, 7), # Frame B
                     c(3, 5:7)) # Frame C
  lapply(pos_select, 
        function(x) sqrt(formula_var(pop_d = pop_d[x],
                                 mu_domains = mu_domains[x],
                                 sd_domains = sd_domains[x])))
}
```

Obtained from the Stack Exchange question: \href{https://stats.stackexchange.com/questions/16608/what-is-the-variance-of-the-weighted-mixture-of-two-gaussians}{What is the variance of the weighted mixture of two gaussians?} (click on the question)

\section{Formula}
Let $X_i$ with $i = 1, ..., n$ be independent random variables each with mean $\mu_i$ and variance $\sigma^2_i$.

Therefore the variance of $Y = \sum_{i=1}^{n} p_i X_i$, with $\sum_i  p_i = 1$ and$\forall p_i>0$ will be
$$Var(Y) = \sum_{i=1}^{n} p_i \sigma_i^2 + \left[\sum_{i=1}^{n} p_i \mu_i^2 
	 - \left(\sum_{i=1}^{n} p_i \mu_i\right)^2 \right]$$

$Y$ is basically a weighted linear combination of independent random variables.

# Toy example

```{r}
# Domain order:
# a - b - c - ab - bc - ac - abc
n_doms <- c(rep(1e3, 3), 1e2, 2e3, 4e2, 1e2) # Population of each domain
mu_s <- c(rep(5, 3), 4, 3, 1, 1) # Means
sd_s <- c(rep(9, 3), 4, 5, 6, 6) # Standard deviations
frame_sd (pop_d = n_doms, mu_domains = mu_s, sd_domains = sd_s)
```

